Paper Wireframe

### 2018-07-16 paper_wireframe_knit.Rmd
Getting tests in logical order and finishing endogenous section

```{r, echo=F}
library('lme4')
library('car')
library('ggplot2')

@@

@@

@@

ggplot(data=dat, aes(y=arrive_breeding))+
  geom_point(aes(x=arrive_uk), col=1, shape=1)+
  geom_point(aes(x=DEPeurope), col=2, shape=1)+
  geom_point(aes(x=DEPnorthAF), col=3, shape=1)+
  geom_point(aes(x=DEPwestAF), col=4, shape=1)+
  geom_point(aes(x=DEPcentralAF), col=5, shape=1)+
  geom_point(aes(x=depart_winterSO), col=6, shape=1)+
  xlim(c(-71, 142))+ylab('Arrival at breeding grounds DOY')+
  xlab('DOY')+theme_classic()

@@

@@

dat$year<-as.character(dat$year)

#reorder breeding location in latitude order

# relevel to central well sampled year
dat$breeding_loc<-factor(dat$breeding, levels=c('Scotland', 'Lake District', 'North York Moors', 'Sherwood Forest',  'Wales', 'Thetford Forest', 'Norfolk Broads','Ashdown Forest', 'New Forest', 'Dartmoor'))

m1<-lmer(DEPwestAF~year+breeding_loc+year+(1|ptt), data=dat)

@@

stargazer(m1, type='text',
covariate.labels=c("Constant: Scotland, 2012","2013", "2014","2015", "2016", "2017", "2018", 'Lake District', 'North York Moors', 'Sherwood Forest',  'Wales', 'Thetford Forest', 'Norfolk Broads','Ashdown Forest', 'New Forest', 'Dartmoor'),
dep.var.labels ="West Africa depature date",
ci=TRUE, intercept.bottom = FALSE, omit.table.layout = "sn", 
digits=1, model.numbers=FALSE,  report='vcs')

@@

env<-read.csv('C:/cuckoo_tracking/data/spring_rainfall_NDVI_GRIMMS_by_stopover_detailcoords_2018_dead.csv', h=T)

@@

dat2<- env %>% group_by(year, ptt, SO_startDOY, variable) %>%
  summarise(country=last(country), SO_endDOY=first(SO_endDOY),
            cuck_pres=first(cuck_pres), precip=mean(value, na.rm=T),
            NDVI= mean(value2, na.rm=T), aquaNDVI=mean(aquaNDVI, na.rm=T),
            aquaANOM=mean(aquaANOM, na.rm=T),terrNDVI=mean(terrNDVI, na.rm=T),
            terrANOM=mean(terrANOM, na.rm=T))

@@

dat2.2<-dat2.1 %>% group_by(year, ptt, SO_startDOY) %>%
  summarise(cumrf=mean(cumrf, na.rm=T),
            NDVI= mean(NDVI, na.rm=T), aquaNDVI=mean(aquaNDVI, na.rm=T),
            aquaANOM=mean(aquaANOM, na.rm=T),terrNDVI=mean(terrNDVI, na.rm=T),
            terrANOM=mean(terrANOM, na.rm=T),
            SOlen=first(SO_endDOY)-first(SO_startDOY))

dat2.3<-dat2.2 %>% group_by(year, ptt) %>%
  summarise(cumrf=last(cumrf),
            NDVI= last(NDVI), 
            aquaNDVI=last(aquaNDVI),
            aquaANOM=last(aquaANOM),
            terrNDVI=last(terrNDVI),
            terrANOM=last(terrANOM),
            last_SOstart=last(SO_startDOY))

>>>>

```{r, echo=F, warning=F, message=F}
library('lme4')
library('car')
library('ggplot2')

@@

#### Data cleaning

@@

## fill NA in mig route

extraz<- read.csv('C:/cuckoo_tracking/data/stopover_bestofday_1daymin_recalc_spring_mig_summary_extras.csv', h=T)

# fills some
dat<-left_join(dat, select(extraz, ptt, year, autumn_mig), by=c('ptt', 'year'))

# fill 2018 birds
dat[c(47, 44, 42),]$autumn_mig.y='SW'
dat[39,]$autumn_mig.y='SE'
# and dead birds
dat[c(2,20,45),]$autumn_mig.y='SW'
dat[c(1,3,23,25,31),]$autumn_mig.y='SE'

names(dat)[names(dat)=='autumn_mig.y']<-'autumn_mig'
dat$autumn_mig.x<-NULL

@@

ggplot(data=dat, aes(y=arrive_breeding))+
  geom_point(aes(x=arrive_uk), col=1, shape=1)+
  geom_point(aes(x=DEPeurope), col=2, shape=1)+
  geom_point(aes(x=DEPnorthAF), col=3, shape=1)+
  geom_point(aes(x=DEPwestAF), col=4, shape=1)+
  geom_point(aes(x=DEPcentralAF), col=5, shape=1)+
  geom_point(aes(x=depart_winterSO), col=6, shape=1)+
  xlim(c(-71, 142))+ylab('Arrival at breeding grounds DOY')+
  xlab('DOY')+theme_bw()

@@

## Endogenous factors that influence dpearture date from West Africa.

No convincing evidence that different breeding populations depart West Africa at different times. *slight significance comes from very early New Forest population*

@@

#reorder breeding location in latitude order
# relevel to central well sampled year
dat$breeding_loc<-factor(dat$breeding, levels=c('Scotland', 'Lake District', 'North York Moors', 'Sherwood Forest',  'Wales', 'Thetford Forest', 'Norfolk Broads','Ashdown Forest', 'New Forest', 'Dartmoor'))

m1<-lmer(DEPwestAF~breeding_loc+(1|year)+(1|ptt), data=dat)

Anova(m1, test.statistic = 'F')

qplot(data=dat, x=breeding_loc, y=DEPwestAF)+theme_bw()+theme(axis.text.x=element_text(angle=90))+xlab("Breeding location")+ylab("Departure date West Africa (DOY)")+geom_label(data=data.frame(), aes(label = 'Anova F=4.55, p=0.023', x = -Inf, y = Inf), hjust = 0, vjust = 1)
```

No evidence that different autumn migration routes impact timing of spring departure from West Africa.

```{r, warning=F, message=F, echo=F}
# and for migratory route
m1<-lmer(DEPwestAF~autumn_mig+(1|year)+(1|ptt), data=dat)

Anova(m1, test.statistic = 'F')

qplot(data=dat, x=autumn_mig, y=DEPwestAF, geom='boxplot')+xlab("Autumn migratory route")+ylab("Departure date West Africa (DOY)")+geom_label(data=data.frame(), aes(label = 'Anova F=0.81, p=0.373', x = -Inf, y = Inf), hjust = 0, vjust = 1)
```

For birds that were tracked through multiple spring migrations there are inter-bird differences in West African departure date and varying degrees of
inter-annual variation within birds.

```{r, warning=F, message=F, echo=F}
multi_yr<-names(which(table(dat$ptt)>1))

multi_yr<-multi_yr[multi_yr!=128300]

m1<-lmer(DEPwestAF~factor(ptt)+(1|year), data=dat[dat$ptt %in% multi_yr ,])

Anova(m1, test.statistic = 'F')

ggplot(data=dat[dat$ptt %in% multi_yr ,], aes(x=factor(ptt), y=DEPwestAF ))+xlab("Bird ID")+ylab("Departure date West Africa (DOY)")+geom_jitter(width=0.08,shape=1)+theme_bw()+geom_label(data=data.frame(), aes(label = 'Anova F=4.13, p=0.017', x = -Inf, y = Inf), hjust = 0, vjust = 1)

## Environmental factors that influence departure date from West Africa.

First, there appears no overall difference between years in departure date from West Africa (median: 102 DOY). *This might support internal regulation for timing of migration out of West Africa by birds e.g. photoperiod (endogenous) but could be an artefact of sample size - more tracking would reveal statistically significant differences between years?*

```{r, warning=F, message=F, echo=F}
dat$year<-as.character(dat$year)

m1<-lmer(DEPwestAF~year+(1|ptt), data=dat)

Anova(m1, test.statistic = 'F')

ggplot(data=dat, aes(x=year, y=DEPwestAF ))+xlab("Year")+ylab("Departure date West Africa (DOY)")+geom_jitter(width=0.08,shape=1)+theme_bw()+geom_label(data=data.frame(), aes(label = 'Anova F=1.27, p=0.315', x = -Inf, y = Inf), hjust = 0, vjust = 1)

@@

stargazer(m1, type='text',
covariate.labels=c("Constant: 2012","2013", "2014","2015", "2016", "2017", "2018"),
dep.var.labels ="West Africa depature date",
ci=TRUE, intercept.bottom = FALSE, omit.table.layout = "sn", 
digits=1, model.numbers=FALSE,  report='vcs')

@@

env<-read.csv('C:/cuckoo_tracking/data/spring_rainfall_NDVI_GRIMMS_TAMSAT_by_stopover_detailcoords_2018_dead.csv', h=T)

@@

dat2<- env %>% group_by(year, ptt, SO_startDOY, variable) %>%
  summarise(country=last(country), SO_endDOY=first(SO_endDOY),
            cuck_pres=first(cuck_pres), precip=mean(value, na.rm=T),
            NDVI= mean(value2, na.rm=T), aquaNDVI=mean(aquaNDVI, na.rm=T),
            aquaANOM=mean(aquaANOM, na.rm=T),terrNDVI=mean(terrNDVI, na.rm=T), terrANOM=mean(terrANOM, na.rm=T), tamRAIN=mean(tamRAIN, na.rm=T), tamRAINanom=mean(tamRAINanom, na.rm=T) )

@@

dat2.2<-dat2.1 %>% group_by(year, ptt, SO_startDOY) %>%
  summarise(cumrf=mean(cumrf, na.rm=T),
            NDVI= mean(NDVI, na.rm=T), aquaNDVI=mean(aquaNDVI, na.rm=T),
            aquaANOM=mean(aquaANOM, na.rm=T),terrNDVI=mean(terrNDVI, na.rm=T),terrANOM=mean(terrANOM, na.rm=T),tamRAIN=mean(tamRAIN, na.rm=T),tamRAINanom=mean(tamRAINanom, na.rm=T), SOlen=first(SO_endDOY)-first(SO_startDOY))

dat2.3<-dat2.2 %>% group_by(year, ptt) %>%
  summarise(cumrf=last(cumrf),
            NDVI= last(NDVI), 
            aquaNDVI=last(aquaNDVI),
            aquaANOM=last(aquaANOM),
            terrNDVI=last(terrNDVI),
            terrANOM=last(terrANOM),
             tamRAIN=last(tamRAIN),
            tamRAINanom=last(tamRAINanom),
            last_SOstart=last(SO_startDOY))
